---
title: "Data Wrangling Part 2:"
subtitle: "Undertanding Your Data: Exercise Solution"
author: "Savet Hong"
date: "`r format(Sys.Date(), '%b %d, %Y')`"
output:
  slidy_presentation:
    incremental: true
---

## Exercises
1. Validate the Column total for each month
2. Transform you clean filing data into a long dataset
    + HINT: it should have three columns: casetype, fiscal_month, filing

## Previously

- 
    ```{r filing, eval=FALSE, echo=TRUE}
library(readxl)
library(tidyverse)    

filing <- read_excel("../data/MockData.xlsx", sheet = "CourtA-Filing", skip = 1)


filing_clean <- filing %>%
      select(- Total) %>%
      rename( Excel_total = X__1) %>%
      mutate(total = rowSums(.[grep("-", names(.))]),
             total_diff = total - Excel_total)
    
View(filing_clean) 
```

## Current Exercise

- Column Validation
    ```{r col_total, eval=FALSE, echo=TRUE}
apply(filing_clean[1:11,-1], 2, sum)

View(filing_clean)        
```    

- Transform the data from wide to long
- 
    ```{r filing_long, echo=TRUE, eval=FALSE}

filing_long <- filing_clean %>%
  #select(-Excel_total, -total,total_diff) %>% # remove explicit columns
  select(`Case Type`:`Jun-18`) %>% # keep range of columns
  filter(`Case Type` != "FILING TOTALS:") %>%
  gather(Fiscal_Month, filing, -`Case Type`)

```

## Exercise 2
- Please read in the Template data and perform the necessary data cleaning.
    + HINT: useful functions: unique(),is.na(), which(), diff()

- What steps needs to be done to clean the data once it's loaded into R?
    + remove blank/irrelevant columns (bullet cells)
    + rename columns
    + add "total" where needed
    + Expand the categories in column A
        - Indentify the unique elements
        - Position between each unique elements
        - Create a vector with the unique vector expanded equal to the distance between elements
        - Repopulate or create a new column with the new vector
    + Remove rows with none relevant data records   

## 

- 
    ```{r template, echo=TRUE, eval=FALSE}
# read in the data
template <-  read_excel("../data/MockData.xlsx", sheet = "Template")  

# check for columns with nas    
sapply(template, function(x) all(is.na(x)))  

##########################   
## clean the imported data
##########################
# remove na columns, rename, and total
template_clean <- template %>%
  select(-X__1, -X__3, -X__4) %>%
  rename(process = CaseType, task = X__2) %>%
  mutate(task = replace(task,is.na(task) & !is.na(`Time per Filing`),"total"))

# Expand Column A as vector
unique_process <- template_clean$process[which(!is.na(template_clean$process))] #unique process (in order)
process_pos <- diff(which(!is.na(template_clean$process))) # position/rows between process
max.pos <- length(unique_process) 
process_vec <- c(rep(unique_process[-max.pos], process_pos),unique_process[max.pos]) # repeat values until next occurance 

# Add expanded vector into dataset
template_clean2 <- template_clean %>%
  mutate(process2 = process_vec) %>%
  select(process2, process, everything()) # a Good place to check your code and data
View(template_clean2)

template_clean2 <- template_clean2 %>%
  filter(!is.na(task)) %>%
  select(-process) %>%
  rename(process = process2)
View(template_clean2)

```
