---
title: "Part 1: Exercises Solution"
author: "Savet Hong"
date: "March 16, 2018"
output:
  slidy_presentation:
  incremental: true
---
  
  
## Part 1: Vector Math
  
- Create object containing a single value

- Create a numerical vector

- Perform arithmatic calculation using these two objects
    + What do you observe?

```{r vec_math, echo=TRUE}
scale <- 3
vec <- c(1:5)

scale*vec
vec-scale

```  

## Key Concept

- Atomic - vector calculation is performed for each element in the vector 
    + imagine the atomic value
        + is calculated on one element at a time across the vector 
        + or it's repeating itself equal to the length of the vector and calculation is then done across the two object 

## Part 2: Matrix Math

- Create an 2x5 matrix object

- Perform various arithmatic calculations using different combination of the three objects
    + What do you observe?

```{r matrix, echo=TRUE} 
mat <- matrix(1:10, 2, 5)
  
scale*mat
vec*mat
mat - vec
  
``` 

## Key Concepts

- Atomic - vector calculation is performed for each element in the vector
- Vector and Matrix calculation is done through repeating values of vector elements equal to the size of the matrix elements

## Arithmatic for Objects of Same Dimension/Size

```{r mat-mat, echo=TRUE}
mat2 <- matrix(11:20, 2,5)
mat2
mat

mat2 - mat

```

## Key Concepts

- Matrix on Matrix calculation (both of equal size) will perform element by element computation
- Matrix on Matrix calculation (not of equal size) will not perform any calculation

## Key Concepts

- Atomic, Vector, and Matrix objects can only contain one mode: numeric, character, and logic
    + e.g., every single element of a matrix must be one of the listed mode above
    + it cannot have one column numeric, another column with character, and another with logic
 
 
## Part 3: Factors

- Create a factor vector

- Change the value of one of the level
    + What happened?
    + How did you achieve this?
    
```{r factors, echo=TRUE}
nums_fact <- factor(rep(c("one","two","three"),5))

levels(nums_fact)

```

```{r eisit_level, echo=TRUE}

nums_fact[nums_fact=="one"] <- "two" 

```

## Add non-exisitng factor level

```{r non-exist-level, echo=TRUE}

levels(nums_fact) <- c(levels(nums_fact), "four")

nums_fact[nums_fact=="three"] <- "four" 

levels(nums_fact)

```  

```{r refactor, echo=TRUE}

nums_fact <- factor(nums_fact)
levels(nums_fact)

```

## Key Concepts

- Factor elements can be changed to another value if a level value already exist
- If the factor value does not exist in the set of level values, then the new level value needs to be added to the existing set prior to updating an exisiting value to the new level value
- Run the factor function on the factor vector to remove levels where value does not exist

