---
title: "Data Wrangling Part 1:"
subtitle: "How to Work with Your Data"
author: "Savet Hong"
date: "`r format(Sys.Date(), '%b %d, %Y')`"
output:
  slidy_presentation:
    incremental: true
---


```{r examples, include=FALSE}
library(readxl)
drinks <- read_excel("../data/drinks_USbirths.xlsx", sheet = "drinks", skip = 2)
births <- read_excel("../data/drinks_USbirths.xlsx", sheet = "USBirths", skip = 2)
```

## Changing the Data
- Index element/object replacement
- Format change/reassignment

- Examples we have done
    ```{r basic_cleaning, eval= FALSE}

nums_fact[nums_fact=="one"] <- "two" 
nums_fact <- factor(nums_fact)
```    

## Handy Package: Tidyverse

- Dplyr (a package within Tidyverse) Common Concepts 
    + mutate: transformation/creation of variable
    + summarise: calculation typically performed on grouped data
    + select: subsetting columns (select specific columns or remove columns)
    + filter: subsetting rows (select specific rows or remove rows)
    + group: associating variables/calculation by classification defined by one or more variables

- Dplyr functions follows the format
    + function(data, operation...)
    + chaining requires 
    data %>%
    function(operation...)

- Tidyr (a package within Tidyverse) Common Concepts 
    + gather: reshape from wide to long
    + spread: reshape from long to wide
    
- Tidyr functions format
    + function(data, key, value, ...)
    + data %>% function(key, value, ...)

## Dplyr Example using the drinks dataset

- Read in the data and call the appropriate libraries

    ```{r drinks, echo= TRUE, eval=FALSE}
library(readxl)
library(tidyverse)

drinks <- read_excel("../data/drinks_USbirths.xlsx", sheet = "drinks", skip = 2)
    
```  

## Example 1
- Create another variable calculating the sum of total serving of alcohol consumed

    ```{r serving, echo=TRUE, eval=FALSE}

# Option 1: R built-in fuctions    
drinks$total_serving = apply(drinks[,2:4], 1, sum)

# Option 2: Dplyr    
drinks <- drinks %>%
  mutate(total_serving = sum(beer_serving:wine_serving))
        
```

## Example 2
- Find the total alcohol consumption for each category.
    ```{r serving_total, echo=TRUE, eval=FALSE}

# Option 1: R built-in fuctions    
total_serving = apply(drinks[,2:ncol(drinks)], 2, sum)

# Option 2: Dplyr    
 drinks %>%
  summarise_if(drinks,is.numeric,sum)
        
```

## Scenario 1: Find the total serving of alcohol for three selected countries.
- Create a subset dataset that includes 3 countries of your choice
    ```{r subset, echo=TRUE, eval=FALSE}
drinks_countries3 <- drinks %>%
  filter(country %in% c("USA", "Canada", "Mexico"))
```

## 
- Reformat the data so that the alcohol type is under the variable "Alcohol_Type" and their associated values under "Serving"
    ```{r long, echo=TRUE, eval=FALSE}
drinks_long <- drinks %>%
      filter(country %in% c("USA", "Canada", "Mexico")) %>%
      gather(key = "Alcohol_Type", value = "Serving", -country)
      
```

## 
- Remove the rows containing "total_litres_of_pure_alcohol" from your data
    ```{r filter21, echo=TRUE, eval=FALSE}
drinks_serving <- drinks_long %>%
  filter(Alcohol_Type != "total_litres_of_pure_alcohol" )
```

- or
    ```{r filter2, echo=TRUE, eval=FALSE}
drinks_serving <- drinks %>%
  filter(country %in% c("USA", "Canada", "Mexico")) %>%
  gather(key = "Alcohol_Type", value = "Serving", -country) %>%
  filter(Alcohol_Type != "total_litres_of_pure_alcohol" )
```


## Concepts Covered
1. Dplyr concepts (those primarily used)
2. Tidyr concepts of data transformation
3. How to use dplyr and tidyr functions


## Reference
- R Reference Card
https://cran.r-project.org/doc/contrib/Short-refcard.pdf

- Data Wrangling Cheat Sheet (Dplyr and Tidyr)
https://www.rstudio.com/wp-content/uploads/2015/02/data-wrangling-cheatsheet.pdf


## Exercises

1. Import the US Birth data
2. Calculate the total births for each month/year
3. Calculate the total births for each year
4. Bonus: Plot your findings.