---
title: "Visualize Your Data"
subtitle: "Graphs"
author: "Savet Hong"
date: "`r format(Sys.Date(), '%b %d, %Y')`"
output:
  slidy_presentation:
  incremental: true
---
  
```{r ex1_msg, echo =TRUE, eval = FALSE, include=FALSE}
## This part of the workshop may require 3 sessions to cover all materials
## Intro + base package:
## ggplot:
## choroplethr: 

```    

## Why create graphs?
    + comprehension
    + understanding
    + analysis
    
## What packages are available?
    + base
    + ggplot
    + choroplethr
    + many more packages available depending on need
    
## Types of Graphs
    + line graph
    + scatter plots
    + bar charts
    + boxplot
    + maps

## Base package: use *plot()* functions
- fuction format
    + **plot(x, y, ...):**
- function require
    + **x:** location of points on the x-axis
    + **y:** location of points on the y-axis
    + **...** optional arguments that can be added to the graph, can include:  
        - plot type (ling, point,steps, etc.)
        - labels for the plot
        - other features

## Example: Base line plot

- 
    ```{r, echo=TRUE, eval=FALSE}
folks <- c("Jane","John","Barry","Bill","Mill","Tom", "Team123")    
```

## base package: use *hist()* functions
- fuction format
    + **hist(x, ...):**
- function require
    + **x:** location of points on the x-axis
    + **...** optional arguments that can be added to the graph, can include:  
        - plot type (ling, point,steps, etc.)
        - labels for the plot
        - other features
 
 ## Example: Base hist

- 
    ```{r, echo=TRUE, eval=FALSE}
   
```

 ## Example: Base boxplot

- 
    ```{r, echo=TRUE, eval=FALSE}
   
```
        
## ggplot package: use *ggplot()* functions
- fuction format
    + **ggplot(data, aes(x,y), ...) +**  geom type, which could be:
      **geom_point()**  
      **geom_hist()**
      **geom_line**  
       and plot type
- function require
    + **data** dataframe to be used
    + **x:** variable within data to be used for location of points on the x-axis
    + **y:** variable within data to be used for location of points on the y-axis
    + **...** optional

 ## Example: ggplot points

- 
    ```{r, echo=TRUE, eval=FALSE}
   
```    

 ## Example: ggplot line

- 
    ```{r, echo=TRUE, eval=FALSE}
   
```


 ## Example: ggplot boxplot

- 
    ```{r, echo=TRUE, eval=FALSE}
   
```


 ## Example: ggplot category

- 
    ```{r, echo=TRUE, eval=FALSE}
   
```

## Choroplethr package: use *county_choropleth()* functions
- fuction format
    + **county_choropleth(data, ...):**
- function require
    + **region:**  geographical location to be mapped
    + **value:** the values to plotted (continuous or categorical) 
    + **...** optional arguments that can be added to the graph, can include:  
        - state_zoom
        - title
        
## Reference

- Base R

- ggplot

-choroplethr

## Concepts covered
-

## Exercises

