---
title: "Data Wrangling Part 4:"
subtitle: "Strings"
author: "Savet Hong"
date: "`r format(Sys.Date(), '%b %d, %Y')`"
output:
  slidy_presentation:
      incremental: true
---
  
```{r, include=FALSE}

```

## Regular Expression
- What is it?


## Commands to Cover
- substr
- sub/gsub
- grep/grepl
- 

## Example 1 : Position

- 
    ```{r ex1_msg, echo =TRUE, eval = FALSE}
 greeting <- "Hello Team 123"
```    

- Extract by position
    + First 5 letters
    + last 3 characters (numbers)
    
- Use *substr()* function
    + requires the text or variable
    + starting position of data element (numeric)
    + ending position of data element (numeric)
    
- 
    ```{r ex1_sub, echo=TRUE, eval=FALSE}
length(greating)
substr(greeting, 1, 5)
substr(greeting)    
```

## Example 2: Pattern
 - Spell out the the word or pattern

```{r, echo=TRUE, eval=FALSE}

```

## Possible Patterns
- Characters
- Numeric
- punctuation

## Example 3: Pattern
- use character class
- grep
- grepl

```{r, echo=TRUE, eval=FALSE}

```

## Example 4: Replacement
- sub
- gsub
```{r, echo=TRUE, eval=FALSE}

```

    ```{r filing_clean, echo=TRUE, eval=FALSE, include= FALSE}
# The grep function is too complicated at this time    
filing_clean <- filing %>%
      select(- Total) %>%
      rename( Excel_total = X__1) %>%
      mutate(total = rowSums(.[grep("-", names(.))]),
             total_diff = total - Excel_total)
    
View(filing_clean)    
```

## Reference

- Regular Expression Cheat Sheet    
https://www.rstudio.com/wp-content/uploads/2016/09/RegExCheatsheet.pdf


## Concepts to be covered
- Character Class
- Locate character match
- Extract by position
- Replace by position
