---
title: "Data Wrangling Part 4:"
subtitle: "Strings: Exercise Solution"
author: "Savet Hong"
date: "`r format(Sys.Date(), '%b %d, %Y')`"
output:
  slidy_presentation:
      incremental: true
---
  

## Exercises
## CourtA-Filing Data
1. Index the Case Type that contains the word "Mental" 
2. Identify and extract Case Type that contains punctuations
3. Create a subset dataframe with columns containing numbers in the variable name

## 
```{r, echo=TRUE, eval=FALSE}
library(tidyverse)
library(readxl)

filing <- read_excel("../data/MockData.xlsx", sheet = "CourtA-Filing", skip = 1)

# Index Casetype with "Mental"
grep("Mental", filing$`Case Type`)

# Extract Casetype with punctuations
filing$`Case Type`[grep("[[:punct:]]", filing$`Case Type`)]

# Subset dataframe with columns contain numbers
filing_sub <- filing[grep("[0-9]", names(filing))]

```


## Bail Data
1. Extract the first two character of the variable Statute and save it to a new variable, "statute_type"
2. Extract all the numerical value in Statue (not in parentheses) and save it to a new variable, "statue_code"
3. Extract all values in parenthese under Statute and save it to a new variable, "statute_sub_code"

## 
```{r, echo=TRUE, eval=FALSE}
bail <- read_excel("../data/MockData.xlsx", sheet = "Bail", na = "NA")

bail_clean <- bail %>%
  mutate(statute_type = substr(Statute, 1,2),
         statute_code = sub("^([A-Z]+)([0-9]+).*", "\\2", Statute),
         statute_sub_code = sub("^([A-Z]+)([0-9]+)", "", Statute))
```

## Address
1. Clean "street" variable in address2 
    a. remove Suites/Apt
    b. remove commas
2. Create variable with only Suite/Apt numbers

## 
```{r,  echo=TRUE, eval=FALSE} 
address <- read_excel("../data/MockData.xlsx", sheet = "Address")       
address2 <- address %>%
      mutate(postal = sub("^([0-9]+) .*", "\\1", Address),
             street = sub("^([0-9]+) (.*)", "\\2", Address),
             street = sub(",","", Address),
             Suite = gsub("[[:alpha:]]","", Address),
             Suite = sub("^([0-9]+) ", "", Suite),
             Suite = gsub("\\s|[[:punct:]]", "", Suite))
```
