---
title: "Data Wrangling Part 4:"
subtitle: "Strings"
author: "Savet Hong"
date: "`r format(Sys.Date(), '%b %d, %Y')`"
output:
  slidy_presentation:
      incremental: true
---
  
```{r, include=FALSE}

```

## Regular Expression
- What is it?
    + sequence of characters 
        - letters, numbers, symbols
    + operations on strings


## Example: Position

- 
    ```{r ex1_msg, echo =TRUE, eval = FALSE}
 greeting <- "Hello Team 123"
```    

- Extract by position
    + First 5 letters
    + last 3 characters (numbers)
    
- Use *substr()* function
    + requires the text or variable
    + starting position of data element (numeric)
    + ending position of data element (numeric)
    + extract the 11 element
    
- 
    ```{r ex1_sub, echo=TRUE, eval=FALSE}
nchar(greeting)
substr(greeting, 1, 5)
substr(greeting, 12, 14)  
substr(greeting, 11, 11)  
```

## What are patterns?
- one or more character that the program tries to match
- Can be...
    + literal (Exact Match)
    + operators (logic, e.g., and/or )
    + constructs (grouping of substring, match part 1, part 2, etc.)


## Use *grep()* or *grepl()* functions
- function require
    + the text or variable
    + pattern to search for
    
- funtion output
    + **grep():** a vector of the element position in original vector that matches the pattern
    + **grepl():** logical vector indicating if each element match the pattern or not

## Example: Exact Pattern
- Spell out the the word or pattern
- 
    ```{r, echo=TRUE, eval=FALSE}
folks <- c("Jane","John","Barry","Bill","Mill","Tom", "Team123")    
grep("T", folks)  
grep("123", folks)

```    

## 

-    
    ```{r, echo=TRUE, eval=FALSE}
library(readxl)
courtb_p1 <- read_excel("../data/MockData.xlsx", sheet = "CourtB-Filing", range = "A1:G13")

var_header <- names(courtb_p1)
grep("-", var_header) # Returns the position of element matching the pattern
grepl("-", var_header) # Returns logic if element matches pattern
```

-  
    ```{r filing_clean, echo=TRUE, eval=FALSE}
library(tidyverse)    

filing <- read_excel("../data/MockData.xlsx", sheet = "CourtA-Filing", skip = 1)

filing_clean <- filing %>%
      select(- Total) %>%
      rename( Excel_total = X__1) %>%
      mutate(total = rowSums(.[grep("-", names(.))]),
             total_diff = total - Excel_total)
    
View(filing_clean)    
```


## Example: Logical Pattern
- Condition in pattern
```{r,  echo=TRUE, eval=FALSE}
grep("B|M", folks)
grep("Civil|Criminal", courtb_p1$`Case Type`)
```

## Non-literal Pattern
- Regular Expression
    + **Characters:** [a-z], [A-Z], [A-z]
    + **Numbers:** [0-9]
    + **Punctuation:** [[:punct:]]

## Special Symbols


## Example 4: Pattern
 - Characters


## Example 5: Pattern
- use regex


```{r, echo=TRUE, eval=FALSE}

```

## Example 4: Replacement
- sub
- gsub
```{r, echo=TRUE, eval=FALSE}

```


## Reference

- Regular Expression Cheat Sheet    
https://www.rstudio.com/wp-content/uploads/2016/09/RegExCheatsheet.pdf


## Concepts to be covered
- Character Class
- Locate character match
- Extract by position
- Replace by position
