---
title: "Data Wrangling Part 4:"
subtitle: "Regular Expression (strings)"
author: "Savet Hong"
date: "`r format(Sys.Date(), '%b %d, %Y')`"
output:
  slidy_presentation:
      incremental: true
---

## Regular Expression
- What is it?
    + sequence of characters 
        - letters, numbers, symbols
        - string objects
    + operations on strings
    
- Types of Operations
    + Extract by position
    + Extract by Pattern 
        - exact
        - pattern
        
## Extract by position

- Use **_substr()_** function
    + **_substr(x, start, stop)_**
    + **x:** the text or variable to apply function on
    + **start:** starting position of data element (numeric) extract
    + **stop:** ending/last position of data element (numeric) extract
    

## Example: Position

- 
    ```{r ex1_msg, echo =TRUE, eval = FALSE}
 greeting <- "Hello Team 123"
```    

- What to extract:
    + First 5 letters
    + last 3 characters (numbers)
    + extract the 11 element
    + hint : use **nchar()** first
    
- 
    ```{r ex1_sub, echo=TRUE, eval=FALSE}
nchar(greeting)
substr(greeting, 1, 5)
substr(greeting, 12, 14)  
substr(greeting, 11, 11)  
```

## Extract by Pattern

- What are patterns?
- One or more character that the program tries to match
- Can be...
    + literal (Exact Match)
    + operators (logic, e.g., and/or )
    + constructs (grouping of substring, match part 1, part 2, etc.)


## Use *grep()* or *grepl()* functions
- fuction format
    + **grep(pattern, x):**
    + **grepl(pattern, x):**
- function require
    + pattern to search for
    + the character (string) vector within which to perform the search

- funtion output
    + **grep():** a vector of the element position in original vector that matches the pattern
    + **grepl():** logical vector indicating if each element match the pattern or not

## Example: Exact Pattern

- 
    ```{r, echo=TRUE, eval=FALSE}
folks <- c("Jane","John","Barry","Bill","Mill","Tom", "Team123")    
```

- Find the element(s) in the vector that contains the following word or pattern:
    + "T"
    + "123"
- 
    ```{r, echo=TRUE, eval=FALSE}
grep("T", folks)  
grep("123", folks)

```    

## Example: Mock Filing Data

-    
    ```{r, echo=TRUE, eval=FALSE}
library(readxl)
library(tidyverse)  
    
courtb_p1 <- read_excel("../data/MockData.xlsx", sheet = "CourtB-Filing", range = "A1:G13")
```

- Identify column/variable names that contains "-"
- 
    ```{r, echo=TRUE, eval=FALSE}
var_header <- names(courtb_p1)
grep("-", var_header) # Returns the position of element matching the pattern
grepl("-", var_header) # Returns logic if element matches pattern
```

- What do you observe in the following?
-  
    ```{r filing_clean, echo=TRUE, eval=FALSE}
  

filing <- read_excel("../data/MockData.xlsx", sheet = "CourtA-Filing", skip = 1)

filing[grep("-", names(filing))]

```

## Notice the difference? 
```{r filing1, echo=TRUE, eval=FALSE}

filing_clean <- filing %>%
      select(- Total) %>%
      rename( Excel_total = X__1) %>%
      mutate(total = rowSums(.[2:13]),
             total_diff = total - Excel_total)

View(filing_clean) 
```


```{r filing2, echo=TRUE, eval=FALSE}
filing_clean <- filing %>%
      select(- Total) %>%
      rename( Excel_total = X__1) %>%
      mutate(total = rowSums(.[grep("-", names(.))]),
             total_diff = total - Excel_total)
    
View(filing_clean)    
```


## Logical Pattern
- Condition in pattern 
- Used with special symbol

## Example: Logical Pattern
- Use the symbol, "|", to specify either or
- In the object, folks, identify:
    + which element starts with a B or  an M?
    + who are they?
- 
    ```{r,  echo=TRUE, eval=FALSE}
grep("B|M", folks)
folks[grep("B|M", folks)]    
```
- In the dataframe, courtb_1, identify:
    + which casetype contains the word Civil or Criminal?
- 
    ```{r,  echo=TRUE, eval=FALSE}    
grep("Civil|Criminal", courtb_p1$`Case Type`)
courtb_p1$`Case Type`[grep("Civil|Criminal", courtb_p1$`Case Type`)]    
```

## Non-literal Pattern

- **Characters:** [a-z], [A-Z], [A-z], [[:alpha:]]
- **Numbers:** [0-9], [[:digit:]]
- **Punctuation:** [[:punct:]]

## Example: Non-literal
- In the object, folks, identify:
    + which element contains characters a-c?
    + which character contains characters e, o, m?
    + who are they?
    
- 
    ```{r,  echo=TRUE, eval=FALSE}
grep("[a-c]", folks)
folks[grep("[a-c]", folks)] 

folks[grep("[eom]", folks)]
```    

- 
    + which element in folks contains numeric value?
    + who are they?
    
- 
    ```{r,  echo=TRUE, eval=FALSE}
grep("[0-9]", folks)
folks[grep("[0-9]", folks)]    
``` 



## Special Matching Symbols

  | either or  
  . any character     
  * match at least 0  
  + match at least 1  
  ? match at most 1 time  
  () grouping

## Anchor Symbols

  ^   start of the string  
  $   end of the string  
  \\b empty string at either end of the word


## Example: Special Symbols

- 
    ```{r, echo=TRUE, eval=FALSE}
address <- read_excel("../data/MockData.xlsx", sheet = "Address")    
```
- Use numerical pattern and anchors to identify elements containing:
    + postal code
    + suites
- Use character pattern to identify elements containing:
    + street names

- 
    ```{r, echo=TRUE, eval=FALSE}
# Identify elements that starts with numerics
grepl("^[0-9]", address$Address)

# Identify elements that ends with numerics    
grepl("[0-9]$", address$Address)

# Identify elements that contains character
grepl("[Z-a]", address$Address)

```

## Replacement

## Use *sub()* or *gsub()* functions
- fuction format
    + **sub(pattern, replacement, x):**
    + **gsub(pattern, replacement, x):**
- function require
    + pattern to search for
    + character to replace match
    + the character (string) vector within which to perform the search

- funtion output
    + **sub():** replace first match in each element of vector, x, that matches the pattern
    + **gsub():** replace all match in each element of vector, x, that matches the pattern

- 
    ```{r, echo=TRUE, eval=FALSE}
folks
sub("l", "BA", folks)
gsub("l", "BA", folks)

```

## Example: Replacement

- Extract and save into new variables:
    + postal address
    + street name
- 
    ```{r,  echo=TRUE, eval=FALSE}    
address2 <- address %>%
      mutate(postal = sub("^([0-9]+) .*", "\\1", Address),
             street = sub("^([0-9]+) (.*)", "\\2", Address))
```
- Note: \\\\1 indicates return first match and \\\\2 indicates return second match

## Reference

- Regular Expression Cheat Sheet    
https://www.rstudio.com/wp-content/uploads/2016/09/RegExCheatsheet.pdf


## Concepts to be covered
- Character Class
- Locate character match
- Extract by position
- Replace by position


## Exercises
### CourtA-Filing Data
1. Index the Case Type that contains the word "Mental" 
2. Identify and extract Case Type that contains punctuations
3. Create a subset dataframe with columns containing numbers in the variable name

### Bail Data
1. Extract the first two character of the variable Statute and save it to a new variable, "statute_type"
2. Extract all the numerical value in Statue (not in parentheses) and save it to a new variable, "statue_code"
3. Extract all values in parenthese under Statute and save it to a new variable, "statute_sub_code"

### Address
1. Clean "street" variable in address2 
    a. remove Suites/Apt
    b. remove commas
2. Create variable with only Suite/Apt numbers